---
title: 'Monthly: VR'
author: "Pitt Sports Science"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Study Information

The study covered in this week's monthly demonstrates the validity of a virtual reality simulation to replicate real-world basketball scenarios and was conducted by Soltani and Morice. The full article can be accessed below:

**Reference**
Soltani P, Morice AHP. A multi-scale analysis of basketball throw in virtual reality for tracking perceptual-motor expertise. Scand J Med Sci Sports. 2023 Feb;33(2):178-188. doi: 10.1111/sms.14250. Epub 2022 Nov 18. PMID: 36315055.

### Data and package import

```{r data and packages, echo = FALSE}
# import packages
library(tidyverse)

# import data
df <- readxl::read_excel('fig4_data_soltani_and_morice_2023.xlsx')

# glimpse at data
head(df)
```

## Including Plots

You can also embed plots, for example:

```{r figure 4, echo=FALSE}
df %>% 
  ggplot(aes(x = velocity, y = angle)) +
  geom_point(aes(shape = type, ))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
