---
title: "02-02 Monthly: Napping"
author: "Pitt Sports Science"
date: "2023-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pitt Sports Science Monthly: Napping

### 02-02 March 2023

This markdown script contains all relevant data to reproduce the figures of the systematic review and meta-analysis by [Mesas et al. 2023](https://pubmed-ncbi-nlm-nih-gov.pitt.idm.oclc.org/36690376/) in Br J Sports Med on the effects of napping on cognitive and physical performance in recreationally active individuals and athletes.

**Reference**
*Mesas AE, Núñez de Arenas-Arroyo S, Martinez-Vizcaino V, Garrido-Miguel M, Fernández-Rodríguez R, Bizzozero-Peroni B, Torres-Costoso AI. Is daytime napping an effective strategy to improve sport-related cognitive and physical performance and reduce fatigue? A systematic review and meta-analysis of randomised controlled trials. Br J Sports Med. 2023 Jan 23:bjsports-2022-106355. doi: 10.1136/bjsports-2022-106355. Epub ahead of print. PMID: 36690376.*


### Import necessary packages

Although there are multiple packages specifically dedicated to generate forest plots common to meta-analyses (e.g., `forestplot`, ), for sake of simplicity and consistency with other monthlys, we recreate these figures using `ggplot2`.

```{r package import, include = TRUE}
library(tidyverse)
```


### Create data

For the monthly, we will only focus on the main result of the study: napping for 30-60 minutes in the afternoon can improve cognitive and physical performance during otherwise normal sleep in recreationally active individuals and athletes.


### Recreate forest plots

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
